Introducing ICBe: An Event Extraction Dataset from Narratives about International Crises

### 2023-07-28 replication_paper/psrm_draft/ICBe_manuscript.tex
pdf and bib

If we could record every international interaction in the realms of
diplomacy, conflict, economics, and beyond, how much unique information
would this chronicle amount to, and how surprised would we be to see
something new? In other words, what is the entropy of international
relations? While this record could in principle be unbounded, the
central conceit of social science is that there are structural
regularities that limit what actors can do, their best options, and even
which actors are likely to survive
(\textbf{brecherInternationalStudiesTwentieth1999?};
\textbf{reiterShouldWeLeave2015?}). If so, then these events can be
recorded and systematically measured by social scientists interested in
these regularities.\footnote{See work on crises
  (\textbf{brecherCrisesWorldPolitics1982?};
  \textbf{beardsleyInternationalCrisisBehavior2020?}), militarized
  disputes (\textbf{palmerMID5Dataset20112021?};
  \textbf{giblerInternationalConflicts181620102018?}), wars
  (\textbf{sarkeesResortWar181620072010?};
  \textbf{reiterRevisedLookInterstate2016?}), organized violence
  (\textbf{ralphsundbergUCDPGEDCodebook2016?};
  \textbf{daviesOrganizedViolence19892022?}), political violence
  (\textbf{raleighIntroducingACLEDArmed2010?}), sanctions
  (\textbf{felbermayrGlobalSanctionsData2020?}), and international
  agreements (\textbf{kinneDefenseCooperationAgreement2020?};
  \textbf{owsiakInternationalBorderAgreements2018?}), dispute resolution
  (\textbf{frederickIssueCorrelatesWar2017?}), and diplomacy
  (\textbf{moyerWhatAreDrivers2020?};
  \textbf{sechserMilitarizedCompellentThreats2011?}).} Thanks to
improvements in natural language processing, more open-ended efforts
have begun to capture entire unstructured streams of events from
international news reports.\footnote{See
  (\textbf{beielerGeneratingPoliticalEvent2016?});
  (\textbf{boscheeICEWSCodedEvent2015?});
  (\textbf{brandtPhoenixRealTimeEvent2018?});
  (\textbf{grantOUEventData2017?});
  (\textbf{liComprehensiveSurveySchemabased2021?}). On event extraction
  from images and social media see
  (\textbf{zhangCASMDeepLearningApproach2019?}) and
  (\textbf{steinert-threlkeldFutureEventData2019?}).} This has invited
fruitful efforts to evaluate the coverage, quality, and accuracy of
attempts to measure international affairs.

We advance existing efforts to identify and structure regularized events
and actor in international politics by combining human coding with
natural language processing to develop (1) a large, flexible ontology of
international affairs and (2) a fine-grained and structured event
dataset of international crises from 1918-2017 developed by applying our
ontology to an unusually high-quality corpus of historical narratives of
international crises
(\textbf{brecherInternationalStudiesTwentieth1999?};
\textbf{wilkenfeldInterstateCrisesViolence2000?};
\textbf{brecherInternationalCrisisBehavior2021?}). We then develop
several methods for objectively gauging how well these event codings
reconstruct the information contained in the original crisis narrative.
We conclude by benchmarking our event codings against several current
state of the art event data collection efforts. The underlying
fine-grained variation in international affairs is unrecognizable
through the lens of current quantification efforts. We find that
existing models produce data on historical episodes that do not contain
enough information to reconstruct the underlying event. In focusing this
initial effort on international crises as a proof of concept sample, we
demonstrate our ontology and method's potential to improve upon existing
empirical identifications of patterns of international interactions.

@@

Much like the recording of chess evolved from natural language
descriptive text notation to the modern figurine algebra notion,
international relations scholars have recently sought to produce a
structured account of historical events by combining the unstructured
corpus of historical records with informative priors about international
relations. The resulting structured account of information about a
historical episode can be combined with that of other events to produce
a systematic account of political events from which we can garner novel
insight, assuming the underlying data have high coverage, precision, and
recall. The easiest way to convey the desired produced of this task is
with an example. (\textbf{fig-cuba-narrative?}) shows a narrative
account of the Cuban Missile Crisis (1962) in natural language sentences
alongside a mapping to discrete machine-readable abstractive events.
From the structured data, scholars can identify similarities and
differences across events concerning important concepts like when
particular foreign policy actions deter versus inflame
(\textbf{jervisCooperationSecurityDilemma1978?};
\textbf{glaserCausesConsequencesArms2000?}), when third parties mediate
in interstate disputes
(\textbf{haffarEmergentPeacemakersCataloguing2002?};
\textbf{quinnPowerPlayMediation2006?}), and how actors try to
communicate resolve (\textbf{tragerDiplomacyWarPeace2016?};
\textbf{luptonReexaminingReputationResolve2018?}). Identifying patterns
of international interactions is not just an inherently interesting
enterprise; it is a necessary precondition to important efforts to
predict where policymakers should turn their attention to improve global
welfare (\textbf{wardLearningSteppingFuture2013?};
\textbf{begerReassessingRoleTheory2021?}).

@@

As we are not the first to attempt to measure international relations in
a structured manner, the columns of Table 1 compare the ontological
coverage of ICBe to existing state of the art systems in production and
with global coverage. We choose these datasets and models as they
represent frequently used and reputable efforts to structure and
describe historical events of interest to scholars of international
politics. The first column starts with our contribution, ICBe, alongside
other event-level datasets including CAMEO dictionary lookup-based
systems (Historical Phoenix
(\textbf{althausClineCenterHistorical2019?}); ICEWS
(\textbf{boscheeICEWSCodedEvent2015?}); Terrier
(\textbf{grantOUEventData2017?})), the Militarized Interstate Disputes
Incidents dataset, and the UCDP-GED dataset
(\textbf{daviesOrganizedViolence19892022?};
\textbf{sundbergIntroducingUCDPGeoreferenced2013?}).\footnote{Other
  related datasets that insufficiently overlap ICBe's domain for
  comparison include BCOW
  (\textbf{lengMilitarizedInterstateCrises1988?}), WEIS
  (\textbf{mcclellandWorldEventInteraction1978?}), CREON
  (\textbf{hermannComparativeResearchEvents1984?}), CASCON
  (\textbf{bloomfieldCASCONIIIComputeraided1989?}), SHERFACS
  (\textbf{shermanSHERFACSCrossParadigmHierarchical2000?}), Real-Time
  Phoenix (\textbf{brandtPhoenixRealTimeEvent2018?}), and COfEE
  (\textbf{balaliCOfEEComprehensiveOntology2021?}) (see histories in
  (\textbf{merrittMeasuringEventsInternational1994?}) and
  (\textbf{schrodtTwentyYearsKansas2006?})).} The final set of columns
compares episode-level datasets beginning with the original ICB project
(\textbf{brecherInternationalCrisisBehavior2021?};
\textbf{brecherCrisesWorldPolitics1982?};
\textbf{beardsleyInternationalCrisisBehavior2020?}); the Militarized
Interstate Disputes dataset (\textbf{palmerMID5Dataset20112021?};
\textbf{giblerInternationalConflicts181620102018?}), and the Correlates
of War (\textbf{sarkeesResortWar181620072010?}). There is imperfect
overlap concerning their intended depth and scope of coverage;
`international crises' are similar, but not identical to, `interstate
wars' and `militarized interstate disputes' which differ yet again from
`individual events of organized violence' and `non-violent action'. Even
like-concepts require care in comparison, as an `aim' in ICBe is the
same as in MIPS, but an `alert' in ICBe is not the same as an `alert' in
MID. Definitions for each ICBe variable are provided in the codebook.

@@

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}m{\dimexpr 0.1in+0\tabcolsep}}{} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{States}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 1.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{(}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\textbf{fazalStateDeathPolitics2011?}}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{)}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{,}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{(}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\textbf{spruytSovereignStateIts1996?}}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{)}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}!{\color[HTML]{000000}\vrule width 2pt}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} \\

@@

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}m{\dimexpr 0.1in+0\tabcolsep}}{} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}m{\dimexpr 2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{Subnational\ Actors}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{(}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\textbf{haffarEmergentPeacemakersCataloguing2002?}}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{)}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}!{\color[HTML]{000000}\vrule width 2pt}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} \\

@@

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}m{\dimexpr 0.1in+0\tabcolsep}}{} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{IGO/NGO}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 1.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{(}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\textbf{bushDensityDeclineFounding2019?}}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{)}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}!{\color[HTML]{000000}\vrule width 2pt}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} \\

@@

The rows in Table 1 represent the types of information we expect to find
in international relations and forms the basis for our proposed
ontology. We create this by performing preliminary natural language
processing of the corpus and identified named entities and behaviors
mentioned in the text. Verbs were matched to the most likely definition
found in Wordnet (\textbf{millerWordNetLexicalDatabase1995?}), tallied,
and then aggregated into a smaller number of hypernyms balancing
conceptual detail and manageable sparsity for human coding (SI Appendix
1.2). This comparison is not intended to fault existing data and models
for not including every variable in ICBe's ontology, as some of these
variables fall outside the scope of a particular dataset's intended
purpose. Rather, it serves as an initial basis for identifying the
heterogeneity in existing efforts to abstract and measure discrete
historical events of interest and to provide theoretical justifications
from existing research about what is included in our dataset's ontology
and where ICBe's detail about historical events can be compared to the
current state of the art.

>>>>

If we could record every international interaction in the realms of
diplomacy, conflict, economics, and beyond, how much unique information
would this chronicle amount to, and how surprised would we be to see
something new? In other words, what is the entropy of international
relations? While this record could in principle be unbounded, the
central conceit of social science is that there are structural
regularities that limit what actors can do, their best options, and even
which actors are likely to survive (Brecher 1999; Reiter 2015). If so,
then these events can be recorded and systematically measured by social
scientists interested in these regularities.\footnote{See work on crises
  (Brecher and Wilkenfeld 1982; Beardsley et al. 2020), militarized
  disputes (Palmer et al. 2021; Gibler 2018), wars (Sarkees and Wayman
  2010; Reiter, Stam, and Horowitz 2016), organized violence (Ralph
  Sundberg and Mihai Croicu 2016; Davies, Pettersson, and Öberg 2022),
  political violence (Raleigh et al. 2010), sanctions (Felbermayr et al.
  2020), and international agreements (Kinne 2020; Owsiak, Cuttner, and
  Buck 2018), dispute resolution (Frederick, Hensel, and Macaulay 2017),
  and diplomacy (Moyer, Turner, and Meisel 2020; Sechser 2011).} Thanks
to improvements in natural language processing, more open-ended efforts
have begun to capture entire unstructured streams of events from
international news reports.\footnote{See Beieler et al. (2016); Boschee
  et al. (2015); Brandt et al. (2018); Grant et al. (2017); Li et al.
  (2021). On event extraction from images and social media see Zhang and
  Pan (2019) and Steinert-Threlkeld (2019).} This has invited fruitful
efforts to evaluate the coverage, quality, and accuracy of attempts to
measure international affairs.

We advance existing efforts to identify and structure regularized events
and actor in international politics by combining human coding with
natural language processing to develop (1) a large, flexible ontology of
international affairs and (2) a fine-grained and structured event
dataset of international crises from 1918-2017 developed by applying our
ontology to an unusually high-quality corpus of historical narratives of
international crises (Brecher 1999; Wilkenfeld and Brecher 2000; Brecher
et al. 2021). We then develop several methods for objectively gauging
how well these event codings reconstruct the information contained in
the original crisis narrative. We conclude by benchmarking our event
codings against several current state of the art event data collection
efforts. The underlying fine-grained variation in international affairs
is unrecognizable through the lens of current quantification efforts. We
find that existing models produce data on historical episodes that do
not contain enough information to reconstruct the underlying event. In
focusing this initial effort on international crises as a proof of
concept sample, we demonstrate our ontology and method's potential to
improve upon existing empirical identifications of patterns of
international interactions.

@@

Much like the recording of chess evolved from natural language
descriptive text notation to the modern figurine algebra notion,
international relations scholars have recently sought to produce a
structured account of historical events by combining the unstructured
corpus of historical records with informative priors about international
relations. The resulting structured account of information about a
historical episode can be combined with that of other events to produce
a systematic account of political events from which we can garner novel
insight, assuming the underlying data have high coverage, precision, and
recall. The easiest way to convey the desired produced of this task is
with an example. (\textbf{fig-cuba-narrative?}) shows a narrative
account of the Cuban Missile Crisis (1962) in natural language sentences
alongside a mapping to discrete machine-readable abstractive events.
From the structured data, scholars can identify similarities and
differences across events concerning important concepts like when
particular foreign policy actions deter versus inflame (Jervis 1978;
Glaser 2000), when third parties mediate in interstate disputes (Haffar
2002; Quinn et al. 2006), and how actors try to communicate resolve
(Trager 2016; Lupton 2018). Identifying patterns of international
interactions is not just an inherently interesting enterprise; it is a
necessary precondition to important efforts to predict where
policymakers should turn their attention to improve global welfare (Ward
et al. 2013; Beger, Morgan, and Ward 2021).

@@

As we are not the first to attempt to measure international relations in
a structured manner, the columns of Table 1 compare the ontological
coverage of ICBe to existing state of the art systems in production and
with global coverage. We choose these datasets and models as they
represent frequently used and reputable efforts to structure and
describe historical events of interest to scholars of international
politics. The first column starts with our contribution, ICBe, alongside
other event-level datasets including CAMEO dictionary lookup-based
systems (Historical Phoenix (Althaus et al. 2019); ICEWS (Boschee et al.
2015); Terrier (Grant et al. 2017)), the Militarized Interstate Disputes
Incidents dataset, and the UCDP-GED dataset (Davies, Pettersson, and
Öberg 2022; Sundberg and Melander 2013).\footnote{Other related datasets
  that insufficiently overlap ICBe's domain for comparison include BCOW
  (Leng and Singer 1988), WEIS (McClelland 1978), CREON (Hermann 1984),
  CASCON (Bloomfield and Moulton 1989), SHERFACS (Sherman 2000),
  Real-Time Phoenix (Brandt et al. 2018), and COfEE (Balali, Asadpour,
  and Jafari 2021) (see histories in Merritt (1994) and Schrodt and Hall
  (2006)).} The final set of columns compares episode-level datasets
beginning with the original ICB project (Brecher et al. 2021; Brecher
and Wilkenfeld 1982; Beardsley et al. 2020); the Militarized Interstate
Disputes dataset (Palmer et al. 2021; Gibler 2018), and the Correlates
of War (Sarkees and Wayman 2010). There is imperfect overlap concerning
their intended depth and scope of coverage; `international crises' are
similar, but not identical to, `interstate wars' and `militarized
interstate disputes' which differ yet again from `individual events of
organized violence' and `non-violent action'. Even like-concepts require
care in comparison, as an `aim' in ICBe is the same as in MIPS, but an
`alert' in ICBe is not the same as an `alert' in MID. Definitions for
each ICBe variable are provided in the codebook.

@@

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}m{\dimexpr 0.1in+0\tabcolsep}}{} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{States}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 1.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{Fazal}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{(2011)}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{,}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{Spruyt}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{(1996)}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}!{\color[HTML]{000000}\vrule width 2pt}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} \\

@@

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}m{\dimexpr 0.1in+0\tabcolsep}}{} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}m{\dimexpr 2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{Subnational\ Actors}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{Haffar}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{(2002)}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.22in+0\tabcolsep}!{\color[HTML]{000000}\vrule width 2pt}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} \\

@@

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 2pt}>{\centering}m{\dimexpr 0.1in+0\tabcolsep}}{} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{IGO/NGO}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 1.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{Bush}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{Hadden}}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{(2019)}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{666666}\vrule width 1pt}>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{✓}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\centering}m{\dimexpr 0.22in+0\tabcolsep}!{\color[HTML]{000000}\vrule width 2pt}}{\textcolor[HTML]{000000}{\fontsize{6}{3}\selectfont{}}} \\

@@

The rows in Table 1 represent the types of information we expect to find
in international relations and forms the basis for our proposed
ontology. We create this by performing preliminary natural language
processing of the corpus and identified named entities and behaviors
mentioned in the text. Verbs were matched to the most likely definition
found in Wordnet (Miller 1995), tallied, and then aggregated into a
smaller number of hypernyms balancing conceptual detail and manageable
sparsity for human coding (SI Appendix 1.2). This comparison is not
intended to fault existing data and models for not including every
variable in ICBe's ontology, as some of these variables fall outside the
scope of a particular dataset's intended purpose. Rather, it serves as
an initial basis for identifying the heterogeneity in existing efforts
to abstract and measure discrete historical events of interest and to
provide theoretical justifications from existing research about what is
included in our dataset's ontology and where ICBe's detail about
historical events can be compared to the current state of the art.

With the exception of large-scale CAMEO dictionary-based systems (the
first grouping of columns), our ontology improves upon the the existing
state of the art quantitative datasets that ignore important information
content about international interactions.\footnote{See Balali, Asadpour,
  and Jafari (2021) for a recent review of ontological depth and
  availability of Gold Standard example text.} We highlight two
particular innovations. First, we separate the `chess pieces' from the
`chess players' in distinguishing between different actors within a
state. By virtue of our ontology coding military versus civilian actors
and national leaders versus bureaucrats, our data can be used to explore
important questions concerning civilian-military relations (Narang and
Talmadge 2017), Track Two diplomacy and the role of sub-national actors
(Hsu et al. 2020), and the evolution of what actors are engaged in
crises - a topic of increasing interest as states engage in gray zone
conflict by employing the coast guard or paramilitary mercenaries
instead of internationally recognized state militaries (J. Andrés Gannon
et al. 2022). Second, we add information about the domains in which
actors behave - whether in land, air, sea, space, or cyber - since they
differ in their technology, tactics, geography, and purpose (Gartzke and
Lindsay 2019). Doing so allows researchers to identify and explain
patterns in escalation conditional on the military means states use in
conflict. Recent concerns about cross-domain conflict and the effect of
new domains of conflict like space and cyber have made this an endeavor
of increased interest to practitioners (J. Andrés Gannon 2022).
